Learning to Synthesize Rules for Network Intrusion Detectors

### 2020-01-17 paper/paper.tex
pass

\begin{figure}[h]
  \vspace{-2ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype}]{pingscan.suricata.synth}
  \vspace{-3ex}  
\end{figure}

@@

\Fix{...listem 5 primeiras regras neste caso...}

\subsection{\Fix{JBoss JMX Console Beanshell Deployer WAR Upload and Deployment Exploit}}
\label{sec:jboss}

@@

\subsubsection{Number of content options.} We empirically observed that content options are very
common. \Luc{Se for preciso temos numeros para fundamentar isso
  melhor, sao cerca de 70K contents nas 28K regras do
  dataset}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
  (\%) das 28K regras tem contents.} As consequence, we defined a
heuristic function that takes the number of content options into
account. We calculate the contribution of number of content options
with the following formula:

@@

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermore, we observed---by analyzing the dataset of
existing rules (see Section~\ref{sec:dataset-benign})---that often
these rules indicate what part of the message the content string
should be checked against.\Mar{Lucas, vc. tem dados para dar suporte a
  isto? ou seja, quantas options do tipo ``content'' possuem
  modificadores.} \Luc{conseguirei esses dados} For example, the
creator of the rule may want to indicate that a specific string be
present in the header of an HTTP request message. Content
\emph{modifiers} serve this purpose. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)
Figure \Fix{xxx}

@@

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''}. The contribution of this option to the score of
the rule that uses it would be $0.5$
(=$\frac{10,111+11}{10,111*2}$). The value of $M_i$ for this case is 2
as the field of the HTTP message \CodeIn{Request URI}---holding the
string ``/HtmlAdaptor''---can be accessed through the modifiers
\CodeIn{http\_uri} and \CodeIn{http\_request\_line}. The relatively
high heuristic value given to this option suggests it is a plausible
option to be used in rules---the string used in the content option
(``/HtmlAdaptor'') is stored in a specific field of the HTTP message
(\CodeIn{Request URI}) and can be accessed by two content modifiers,
namely (\CodeIn{http\_uri} and \CodeIn{http\_request\_line}). For the
option \CodeIn{content: ``keep-alive''}, either the modifier
\CodeIn{http\_header} or the modifier \CodeIn{http\_connection} could
be used to locate the string ``keep-alive'' stored in the field
\CodeIn{Connection} (see Figure~\Fix{xxx}). Consequently, for this
case, the contribution of modifiers to this content option would be
$0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
used in a rule, the aggregated contribution of this heuristic function
would be 0.35 (=(0.2+0.5)/2).

@@

\vspace{1ex}
\noindent
\textbf{Malicious traffic.}~Nikito?\Fix{where do these
  attacks/messages come from?}

>>>>

\begin{figure}[h]
  \vspace{-2ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{pingscan.suricata.synth}
  \vspace{-3ex}  
\end{figure}

@@

\begin{figure}[h]
  \vspace{-2ex}
  \lstinputlisting[language=C,numbers=none,frame=none,keywords={dsize,icode,itype,icmp\_id,icmp\_seq}]{top-pingscan.rules}
  \vspace{-3ex}  
\end{figure}

\Luc{Apenas 3 regras tiveram f1 de 100\%. O total de regras nesse caso foi de 13.}
\Fix{...listem 5 primeiras regras neste caso...}

\subsection{Remote Code Execution}
\label{sec:rce}

"RCE refers to the process by which an agent can exploit a network vulnerability to run arbitrary code on a targeted machine or system. In an RCE attack, hackers intentionally exploit a remote code execution vulnerability to run malware. By prompting the targeted device to perform code execution, a hacker can run their own programming in its place. This programming can then enable them to gain full access, steal data, carry out a full distributed denial of service (DDoS) attack, destroy files and infrastructure, or engage in illegal activity. And as the term remote execution suggests, an RCE cyberattack can take place from any geophysical location."\\ Texto copiado de:  https://www.solarwindsmsp.com/blog/remote-code-execution
\Luc{---------------- estou aqui}

@@

\subsubsection{Number of content options.} We empirically observed that content options are very
common. \Luc{Se for preciso temos numeros para fundamentar isso
  melhor, sao cerca de 70K contents nas 28K regras do
  dataset}\Mar{sim, isto ajuda, porem mais importante eh saber quantas
  (\%) das 28K regras tem contents.} \Luc{sao 69417 contents nas 26094 regras q utilizam contents} As consequence, we defined a
heuristic function that takes the number of content options into
account. We calculate the contribution of number of content options
with the following formula:

@@

\subsubsection{Frequency of content modifiers.} As previously
mentioned, \percRulesWithContent\ of Suricata rules involve content
options. Furthermore, we observed---by analyzing the dataset of
existing rules (see Section~\ref{sec:dataset-benign})---that often
these rules indicate what part of the message the content string
should be checked against.\Mar{Lucas, vc. tem dados para dar suporte a
  isto? ou seja, quantas options do tipo ``content'' possuem
  modificadores.} \Luc{das 26k regras que utilizam content, 13k delas utilizam modificadores http q sao os q nós consideramos.} For example, the
creator of the rule may want to indicate that a specific string be
present in the header of an HTTP request message. Content
\emph{modifiers} serve this purpose. For example, the modifier
\CodeIn{http\_header} in the option \CodeIn{content: ``keep-alive'';
  http\_header} states the requirement that the string ``keep-alive''
be present in the header of the message. (A keep-alive connection
attribute indicates that a single TCP connection should remain open
for multiple HTTP requests/responses.)
Figure~\ref{fig:http-header-example}

\begin{figure}[h!]
\centering
\includegraphics[scale=0.5]{figs/http_header-example.png}
\caption{\CodeIn{http\_header} example.}
\label{fig:http-header-example}
\end{figure}

@@

Consider a rule containing the option \CodeIn{content:
  ``/HtmlAdaptor''}. The contribution of this option to the score of
the rule that uses it would be $0.5$
(=$\frac{10,111+11}{10,111*2}$). The value of $M_i$ for this case is 2
as the field of the HTTP message \CodeIn{Request URI}---holding the
string ``/HtmlAdaptor''---can be accessed through the modifiers
\CodeIn{http\_uri} and \CodeIn{http\_request\_line}. The relatively
high heuristic value given to this option suggests it is a plausible
option to be used in rules---the string used in the content option
(``/HtmlAdaptor'') is stored in a specific field of the HTTP message
(\CodeIn{Request URI}) and can be accessed by two content modifiers,
namely (\CodeIn{http\_uri} and \CodeIn{http\_request\_line}). For the
option \CodeIn{content: ``keep-alive''}, either the modifier
\CodeIn{http\_header} or the modifier \CodeIn{http\_connection} could
be used to locate the string ``keep-alive'' stored in the field
\CodeIn{Connection} (see Figurgure~\ref{fig:http-header-example}). Consequently, for this
case, the contribution of modifiers to this content option would be
$0.2$ (=$\frac{4089+29}{10111*2}$). Assuming that both options are
used in a rule, the aggregated contribution of this heuristic function
would be 0.35 (=(0.2+0.5)/2).

@@

\vspace{1ex}
\noindent
\textbf{Malicious traffic.}~Nikto?\Fix{where do these
  attacks/messages come from?}
  \Luc{Nós executamos um servidor http (Apache2) numa maquina e executamos o Nikto em outra maquina atacando o servidor http. Capturamos o trafego q chegava ao servidor e eh ele q estamos utilizando.}
